Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-04-08 SIR_manuscript.tex
simplify V^-1 to a user friendly one
more on d R0/d rho, remove old analysis
more Appendix

where parameters are specified in Table \ref{tab:params}. The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible. 
Matrices $F$ and $V$ are \ali{all $S\_u \& S\_n$ should be $S\_u^* \& S\_n^*$}
\begin{align}
\label{FV}
F =& \beta/N_0 \left[ \begin {array}{cccc} 
S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_c)\,S\_u\\
S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_c)\,S\_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right], \\
  V =&
 \left[ \begin {array}{cccc}  
F_I+\gamma&-\omega&0&0\\
-(1-p_I)F_I&\omega+\gamma&0&0\\
-p_I F_I&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right], \text{thus}\\
V^{-1} =&
\left[ \begin {array}{cccc}
\frac {\omega+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&\frac {\omega}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
\noalign{\medskip}
\frac{(1-p_I) F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
\frac{F_I+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
\noalign{\medskip}
\frac{p_I F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
\frac{\omega p_I F_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma)}& \frac{1}{\omega+\gamma}&0 \\
\noalign{\medskip}
\frac{\omega\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I) \gamma}& 
\frac{\omega^2\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma) \gamma}&
\frac{\omega}{(\omega+\gamma) \gamma}&
\frac{1}{\gamma}
\end {array} \right].
\end{align}

The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
\begin{equation}
\label{mat:G}
G = \left[ \begin {array}{cc}
G_{11}&G_{12}\\
0&0
\end {array} \right], \text{ where } \\
G_{11} = \frac{\beta}{\gamma}\ C
\left[\begin {array}{cc}
A\,S\_u & B\,S\_u\\
A\,S\_n & B\,S\_n
\end {array}\right],
\end{equation}
with expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
Also note that, all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.

>>>>

where parameters are specified in Table \ref{tab:params}. The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible. 
Matrices $F$ and $V$ are \ali{all $S\_u \& S\_n$ should be $S\_u^* \& S\_n^*$}
\begin{align}
\label{FV}
F =& \beta/N_0 \left[ \begin {array}{cccc} 
S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_c)\,S\_u^*\\
S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_c)\,S\_n^* \\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right] \\
 =&\beta\left[ \begin {array}{c} x\_n^* \\ x\_u^* \\ 0 \\ 0 \end {array} \right]
        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
 \left[ \begin {array}{cccc}  
F_I+\gamma&-\omega&0&0\\
-(1-p_I)F_I&\omega+\gamma&0&0\\
-p_I F_I&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right].
\end{align}
The matrix inverse of $V$ is 
\begin{align}
\label{vinv}
V^{-1} =&
\frac{1}{\gamma C}
\left[ \begin {array}{cccc}
\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
\gamma(\omega+\gamma)(1-p_I)F_I & \gamma(\omega+\gamma)(F_I+\gamma)&0&0 \\ \noalign{\medskip}
\gamma(\omega+\gamma) p_I F_I & \gamma\omega p_I F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
\omega(\omega+\gamma) p_I F_I & \omega^2 p_I F_I & C \omega/(\omega+\gamma) & C
\end {array} \right],
\end{align}
where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)F_I \big) (\omega+\gamma)$. Note that all the columns of matrix $V^{-1}$ summ up to $C$.

The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as the following block matrix.
\begin{equation}
\label{mat:G}
G = \left[ \begin {array}{cc}
G_{11}&G_{12}\\
0&0
\end {array} \right],
\end{equation}
where the 2 by 2 submatrix $G_{11}$ is 
\begin{equation}
\label{G11}
G_{11} = \frac{\beta}{\gamma \, C} 
\left[ \begin {array}{c}  x\_n^* \\ x\_u^* \\ 0 \\ 0  \end {array} \right]
\left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
\left[ \begin {array}{cc}
\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
\gamma(\omega+\gamma)(1-p_I)F_I & \gamma(\omega+\gamma)(F_I+\gamma) \\ \noalign{\medskip}
\gamma(\omega+\gamma) p_I F_I & \gamma\omega p_I F_I \\ \noalign{\medskip}
\omega(\omega+\gamma) p_I F_I & \omega^2 p_I F_I
\end {array} \right].
\end{equation}

It is notable that matrix $F$ \eqref{FV} and consequently $G_{11}$ have rank one. That is $G_{11}$ has only one non-zero eigenvalue which we refer to it the basic reproduction number $\Rnum$ (defined as the spectral radius of $G$). Also the block matrix $G_{12}$ does not influence $\Rnum$ and all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.

$\Rnum$ has a complicated expression and depend on the type of analysis, the expression can be simplifed and presented accordingly. For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $A$, $B$ and $C$, specified in \eqref{eq:abc}. This is intended to show that $\Rnum$ decreases as the strength of isolation, shown by $\theta\_w$ and $\theta\_c$, increases.

The hard thing to prove is to show the reduction of $\Rnum$ with respect to testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$.
